Realizing temporal graphs from fastest travel times

### 2023-02-09 ICALP23-shortversion.tex
Final versions_001

\usepackage{thmtools} 
\usepackage{thm-restate}
\usepackage{cite}

\usepackage{caption} %for subfigure - join multiple figures and add captions
\usepackage{subcaption}
\usepackage{changepage} %inside figure we can put \begin{adjustwidth}{-1cm}{-1cm} \end{adjustwidth} and we can ignore the margins
\usepackage{mathtools} % write text [under]{over} arrow $\xrightarrow[\text{world}]{\text{hello}}$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PROBLEMS NAMES
%% One label per edge, period Delta, exact realization
\newcommand{\deltaExactLong}{\textsc{Periodic Temporal Graph Realization}}
\newcommand{\deltaExact}{\textsc{Periodic TGR}}

@@

\bibliographystyle{plainurl}% the mandatory bibstyle
\usepackage{cite} %order citations by increasing number

\title{Realizing temporal graphs from fastest travel times} %TODO Please add

@@

%Due to space constraints, proofs of results marked with $\star$ are (partially) deferred to Appendix.

@@

\subparagraph{Preliminaries and notation.}
We already introduced the most central notion and concepts. There are some additional definitions we need to present our proofs and results which we give in the following.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.
We denote with $d(u,v,P)$ the duration of the fastest temporal path from $u$ to $v$ along path $P$. Note that $d(u,v,P) \geq d(u,v)$.

@@

\subparagraph{Organization of the paper.}
In \cref{sec:hardness} we present our hardness results, first the NP-hardness in \cref{sec:nphardness} and then the parameterized hardness in \cref{sec:w1hardness}. In \cref{sec:algos} we present our algorithmic results. First we give in \cref{sec:cyclealgo} a polynomial-time algorithm for the case where the underlying graph is a tree. In \cref{sec:FPT} we generalize this and present our FPT result, which is the main result in the paper. In \cref{sec:cyclealgo} we provide a simple polynomial-time algorithm for the case where the underlying graph is a cycle. Finally, we conclude in \cref{sec:conclusion} and discuss some future work directions.

\section{Hardness results for \deltaExact}\label{sec:hardness}
In this section we present our main computational hardness results. In \cref{sec:nphardness} we show that \deltaExact\ is NP-hard even for constant $\Delta$. In \cref{sec:w1hardness} we investigate the parameterized computational hardness of \deltaExact\ with respect to structural parameters of the underlying graph. We show that \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.

\subsection{NP-hardness of \deltaExact}\label{sec:nphardness}
In this section we prove that in general it is NP-hard to determine a $\Delta$-periodic temporal graph $(G,\lambda)$ respecting a duration matrix $D$,
even if $\Delta$ is a small constant.

\begin{theorem}\label{thm:NPhardness}
	\deltaExactLong\ is \NP-hard for all $\Delta \geq 3$.
\end{theorem}

@@

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
\begin{tikzpicture}%[xscale=2]

@@

\end{tikzpicture}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(\overline{x}_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=\Delta$, $d(c,x_i)=2\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

$(\Rightarrow)$: Assume the constructed instance $(D,\Delta)$ of \deltaExact\ is a \textsc{yes}-instance. 
Then there exist a label $\lambda(e)$ for each edge $e\in E(G)$ such that for each vertex pair $u,w$ in the temporal graph $(G,\lambda,\Delta)$ we have that a fastest temporal path from $u$ to $w$ is of duration $d(u,w)$. 
%In particular, for any two vertices $u,w$ with $d(u,w)=\infty$ we have that there does not exist a temporal path from $u$ to $w$ in $(G,\lambda)$.

We construct a satisfying assignment for $\phi$ as follows. For each variable $x_i$, 
if $\lambda(x_i x_i^T)=\lambda(x_i^T v)$, then we set $x_i$ to \textsc{true}, otherwise we set $x_i$ to \textsc{false}.

To show that this yields a satisfying assignment, we need to prove some properties of the labeling $\lambda$.
First, observe that adding an integer $t$ to all time labels does not change the duration of any temporal paths. 
Second, observe that if for two vertices $u,w$ we have that $d(u,w)$ equals the distance between $u$ and $w$ in $G$
(\ie the duration of the fastest temporal path from $u$ to $w$ equals the distance of the shortest path between $u$ and $w$), 
then there is a shortest path $P$ from $u$ to $w$ in $G$ such that 
the labeling $\lambda$ assigns consecutive time labels to the edges of $P$.

Let $\lambda(x_i x_i^T)=t$ and $\lambda(x_i x_i^F)=t'$, for an arbitrary variable $x_i$. 
If both $\lambda(x_i^T v)\neq t+1$ and $\lambda(x_i^F v)\neq t'+1$, then $d(x_i,v)>2$, which is a contradiction. 
Thus, for every variable $x_i$, we have that $\lambda(x_i^T v)= t+1$ or $\lambda(x_i^F v)= t'+1$ (or both). 
In particular, this means that if $\lambda(x_i x_i^F)=\lambda(x_i^F v)$, then we set $x_i$ to \textsc{false}, since in this case $\lambda(x_i x_i^T)\neq\lambda(x_i^T v)$.

%Furthermore, we have that if $x$ appears non-negated in clause $c$, then $\lambda(\{c,x^T\})=t-1$, otherwise we have $d(c,x^T)>2$. Symmetrically, we have that if $x$ appears negated in clause $c$, then $\lambda(\{c,x^F\})=t'-1$, otherwise we have $d(c,x^F)>2$.

Now assume for a contradiction that the described assignment is not satisfying. Then there exists a clause $c$ that is not satisfied. 
Suppose that $x_1, x_2, x_3$ are three variables that appear in $c$.
Recall that we require $d(c,v)=2$ and $d(v,c)=\Delta -1$. 
The fact that $d(c,v)=2$ implies that we must have a temporal path consisting of two edges from $c$ to $v$, 
such that the two edges have consecutive labels. 
By construction of $G$ there are three candidates for such a path, one for each literal of $c$. 
Assume w.l.o.g.\ that $x_1$ appears in $c$ non-negated (the case of a negated appearance of $x_1$ is symmetrical) and that the temporal path realizing $d(c,v)=2$ goes through vertex $x_1^T$. 
Let us denote with $t = \lambda(x_1^T v)$.
It follows that $\lambda(x_1^T c)=\lambda(x_1^T v)-1 = t - 1$.
Furthermore, since $d(c,x_1)=2$ we also have that $\lambda(x_1^T c)=\lambda(x_1 x_1^T)-1$. 
Therefore $\lambda(x_1 x_1^T)=\lambda(x_1^T v) = t$. 
Which implies that $x_1$ is set to \textsc{true}.
%
Let us observe paths from $v$ to $c_1$.
We know that $d(v,c)=\Delta -1$.
The underlying path of the fastest temporal path from $v$ to $c$, that goes through $x_1^T$ is the path $P = (v,x_1^T,c)$.
Since $\lambda(x_1^Tc) > \lambda(x_1^T v)$ we get that the duration of the temporal path $(P,\lambda)$ is equal to 
$d(P,\lambda)= (\Delta + t-1) - t + 1 = \Delta$.
This implies that
the fastest temporal path from $v$ to $c$ is not $(P,\lambda)$ and therefore does not pass through $x_1^T$.
Since there are only two other vertices connected to $c$, 
we have only two other edges incident to $c$, that can be used on a fastest temporal path
$v$ to $c$.
Suppose now w.l.o.g.\ that also $x_2$ appears in $c$ non-negated (the case of a negated appearance of $x_2$ is symmetrical) and that the temporal path realizing $d(v,c)=\Delta-1$ goes through vertex $x_2^T$.
Let us denote with $t' = \lambda(x_2^T v)$.
Since the fastest temporal path from $v$ to $c$ is of the duration $\Delta - 1$,
and the edge $x_2^T c$ is the only edge incident to vertex $c$ and edge $ x_2^T v$,
it follows that $\lambda(x_2^T c) \geq \lambda(x_2^T v) - 2 = t' - 2$.
Since $d(x_2,v) = 2$ it follows that $\lambda(x_2 x_2^T) = \lambda (x_2^T v) - 1 = t' - 1$.
Knowing this and the fact that $d(x_2,c)=2$, we get that $\lambda(x_2^T c)$ must be equal to $t'-2$.
Therefore the fastest temporal path from $v$ to $c$ passes through edges $ x_2^Tv$ and $x_2^Tc$.
In the above we have also determined that $\lambda(x_2x_2^T) \neq \lambda ( x_2^Tv)$,
which implies that $x_2$ is set to \textsc{false}.
But now we have that $x_1,x_2$ both appear in $c$ non-negated, where one of them is \textsc{true}, while the other is \textsc{false},
which implies that the clause $c$ is satisfied, a contradiction.

%Assume w.l.o.g.\ that variable $x$ appears non-negated in clause $c$. Then we have that $\{c,x^T\}\in E(G)$. Assume that $\lambda(\{c,x^T\})=t$. Recall that we have specified $d(c,x_2)=2$. Hence we have that $\lambda(\{x_2,x^T\})=t+1$.

$(\Leftarrow)$: Assume that $\phi$ is satisfiable. Then there exists a satisfying assignment for the variables in $\phi$.

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{true}, we set $\lambda(x_i^F v)=3$.
    \item If variable $x_i$ is set to \textsc{false}, we set $\lambda(x_i^T v)=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all duratios are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

    If $x_i$ is set to \textsc{true}, then there is a temporal path from $x_i$ to $v$ via 
    $x_i^F$ of duration $2$, since
    $\lambda(x_i x_i^F)=2$ and $\lambda(x_i^F v)=3$.
    For a temporal path from $v$ to $x_i$ we observe the following.
    The only possible labels to leave the vertex $v$ are $2$ and $3$, which take us from $v$ to $x_j^T$ or $x_j^F$ of some variable $x_j$.
    The only two edges incident to $x_i$ have labels $2$, therefore the fastest path from $v$ to $x_i$
    cannot finish before the time $\Delta + 2$.
    The fastest way to leave $v$ and enter to $x_i$ would then be to leave $v$ at edge $x_i^Fv$ with label $3$,
    and continue to $x_i$ at time $\Delta + 2$,
    which gives us the desired duration $\Delta$.

    If $x_i$ is set to \textsc{false}, then, by similar arguing, 
    there is a temporal path from $x_i$ to $v$ via $x_i^T$ of duration $2$,
    and a temporal path from $v$ to $x_i$, through $x_i^F$ of duration $\Delta$.

    \item For each clause $c$ in $\phi$ we have to check that $d(c,v)=2$
    and $d(v,c)=\Delta - 1$:

    Suppose $x_i,x_j,x_k$ appear in $c$.
    Since we have a satisfying assignment at least one of the literals in $c$ is set to \textsc{true} and at least one to \textsc{false}. 
    Suppose $x_i$ is the variable of the literal that is \textsc{true} in $c$,
    and $x_j$ is the variable of the literal that is \textsc{false} in $c$.
    Let $x_i$ appear non-negated in $c$ and is therefore set to \textsc{true} (the case when $x_i$ appears negated in $c$ and is set to \textsc{false} is symmetric).
    Then there is a temporal path from $c$ to $v$ through $x_i^T$ such that $\lambda(x_i^T c)=1$ and $\lambda(x_i^T v)=2$. 
    Let $x_j$ appear non-negated in $c$ and is therefore set to \textsc{false} 
    (the case when $x_j$ appears negated in $c$ and is set to \textsc{true} is symmetric).
    Then there is a temporal path from $v$ to $c$ through $x_j^T$ such that $\lambda(x_j^T v)=3$ and $\lambda(x_j^T c)=1$,
    which results in a temporal path from $v$ to $c$ of duration $\Delta -1$.

    \item Let $x_i$ be a variable that appears in clause $c$.
    If $x_i$ appears non-negated in $c$ we have to check that $d(c,x_i)=d(c,x_i^F)=2$
    and $d(x_i,c)=d(x_i^F,c)= \Delta$.

    There is a temporal path from $c$ to $x_i$ via $x_i^T$ and also a temporal path from $c$ to $x_i^F$ via $x_i^T$ such that $\lambda(x_i^T c)=1$ and $\lambda(x_i x_i^T)=\lambda(x_i^T x_i^F)=2$,
    which proves the first equality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^T$ and
    second, from $x_i^F$ to $c$ through $x_i^T$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(x_i x_i^T) = \lambda (x_i^T x_i^F)=2$ and 
    finish on the edge with label $1$, as $\lambda(x_i^T c) = 1$.

    If $x$ appears negated in $c$ we have to check that 
    $d(c,x_i)=d(c,x_i^T)=2$
    and $d(x_i,c)=d(x_i^T,c)= \Delta$.

    There is a temporal path from $c$ to $x$ via $x^F$ and also a temporal path from $c$ to $x^T$ via $x^F$ such that $\lambda(c x^F)=1$ and $\lambda(x x^F)=\lambda(x^T x^F)=2$,
    which proves the first inequality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^F$ and
    second, from $x_i^T$ to $c$ through $x_i^F$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(x_i x_i^F) = \lambda (x_i^T x_i^F)=2$ and 
    finish on the edge with label $1$, as $\lambda(x_i^F c) = 1$.
    Which proves the second equality.

    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we have to check that
    first,
    $d(c,x_i^T)=d(c,x_i^F)=2$,
    second, $d(x_i^T, c) = d(x_i^F,c)=\Delta$,
    third, $d(x_i,c)=\Delta$,
    and fourth $d(c,x_i)=2 \Delta + 2$.

    Let $x_j$ be a variable that appears non-negated in $c$ (the case where $x_j$ appears negated is symmetric). 
    Then there is a temporal path 
    from $c$ to $x_i^T$ via $x_j^T$ and also a temporal path 
    from $c$ to $x_i^F$ via $x_j^T$ such that $\lambda(x_j^T c)=1$ and $\lambda(x_j^T x_i^T)=\lambda(x_j^T x_i^F)=2$,
    which proves the first equality.
    Using the same temporal path in the opposite direction,
    \ie first the edge $x_j^T c$ and then one of the edges $x_j^T x_i^F$ or $x_j^T x_i^T$,
    yields the second equality.
    For a temporal path from $c$ to $x_i$ we traverse the following three edges 
    $x_j^T c, x_j^T x_i^F$ and $x_i^Fx_i$,
    with labels $1, 2, 2$ respectively.
    This proves the second equality.
    Now for the case of a temporal path from $x_i$ to $c$,
    we use the same three edges, but in the opposite direction,
    namely $x_i^Fx_i, x_j^T x_i^F$ and $x_j^T c$,
    again with labels $2, 2, 1$, respectively,
    which proves the last equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we have to check that
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.

    There is a path from $x_i$ to $x_j$ that passes first through one of the vertices 
    $x_i^T$ or $x_i^F$, 
    and then through one of the vertices $x_j^T$ or $x_j^F$.
    This temporal path is of length $3$, where all of the edges have label $2$,
    which proves the first equality.
    Now, a temporal path from $x_i$ to $x_j^T$ (resp.~$x_j^F$),
    passes through one of the vertices $x_i^T$ or $x_i^F$.
    This path is of length two, where all of the edges have label $2$,
    which proves the second equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we have to check that
    $d(c_i,c_j)= \Delta + 1$.

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$, respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, since we have $\Delta\ge 3$.
\end{proof}

\subsection{Parameterized hardness of \deltaExact}\label{sec:w1hardness}

In this section, we investigate the parameterized hardness of \deltaExact\ with respect to structural parameters of the underlying graph. We show that the problem is W[1]-hard when parameterized by the feedback vertex number of the underlying graph. The \emph{feedback vertex number} of a graph $G$ is the cardinality of a minimum vertex set $X\subseteq V(G)$ such that $G-X$ is a forest. The set $X$ is called a \emph{feedback vertex set}. 
Note that, in contrast to the result of the previous section (\cref{thm:NPhardness}), the reduction we use to obtain the following result does not produce instances with a constant $\Delta$.

\begin{theorem}\label{thm:W1wrtFVS}
    \deltaExactLong\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExactLong\ where the period $\Delta$ is infinite (that is, there are no periods) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. At the end of the proof we show how to obtain the result for a finite $\Delta$ and a matrix $D$ of durations of fastest paths, that only has finite entries.

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExactLong\ (with infinity entries and no periods) as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to be an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.
    We will construct $G$ using several gadgets.

    \subparagraph{Edge selection gadget.} We first introduce an \emph{edge selection gadget $G_{i,j}$ for color combination $i,j$} with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $4n+1$ vertices each, that is, $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, 
        %u^\ell_n, u^\ell_{n+1},\ldots, u^\ell_{4n-1}, 
        u^\ell_{4n}\}$ for all $\ell\in[m]$. 
        %Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{4n},v_{i,j}^{\star\star}\}$.

>>>>

\usepackage{thmtools} 
\usepackage{thm-restate}

\usepackage{caption} %for subfigure - join multiple figures and add captions
%\usepackage{subcaption}
\usepackage{changepage} %inside figure we can put \begin{adjustwidth}{-1cm}{-1cm} \end{adjustwidth} and we can ignore the margins
\usepackage{mathtools} % write text [under]{over} arrow $\xrightarrow[\text{world}]{\text{hello}}$

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PROBLEMS NAMES
%% One label per edge, period Delta, exact realization
\newcommand{\deltaExactLong}{\textsc{Simple periodic Temporal Graph Realization}}
\newcommand{\deltaExact}{\textsc{Simple TGR}}

@@

\bibliographystyle{plainurl}% the mandatory bibstyle
\usepackage{cite} %order citations by increasing number
\title{Realizing temporal graphs from fastest travel times} %TODO Please add

@@

Due to space constraints, proofs of results marked with $\star$ are (partially) deferred to the Appendix.

@@

\subparagraph{Preliminaries and notation.}
We already introduced the most central notion and concepts. There are some additional definitions we need, to present our proofs and results which we give in the following.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$. 
For simplicity of the presentation~--and with a slight abuse of notation--~we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ whenever it is clear from the context.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda)$,
where we actually mean that $(P,\lambda,\Delta)$ is a temporal path with $P$ as its underlying (static) path.

@@

\section{Hardness results for \deltaExact}  
In this section we present our main computational hardness results.  
We first show that \deltaExact\ is NP-hard even for constant $\Delta$. 
%Then we investigate the parameterized computational hardness of \deltaExact\ with respect to structural parameters of the underlying graph. We show that \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.

%\subparagraph{NP-hardness of \deltaExact.}\label{sec:nphardness}
%We prove that in general it is NP-hard to determine a $\Delta$-periodic temporal graph $(G,\lambda)$ respecting a duration matrix $D$,
%even if $\Delta$ is a small constant.

\begin{theorem}[$\star$]\label{thm:NPhardness}
	\deltaExact\ is \NP-hard for all $\Delta \geq 3$.
\end{theorem}

@@

%
\begin{figure}[htb]
	\noindent
	\makebox[\textwidth]{
		\centering
\begin{tikzpicture}%[xscale=2]

@@

\end{tikzpicture}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(\overline{x}_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=\Delta$, $d(c,x_i)=2\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. We defer the correctness proof to the appendix.
\end{proof}

\end{comment}

%\subparagraph{Parameterized Hardness of \deltaExact.}\label{sec:w1hardness}

Next, we investigate the parameterized hardness of \deltaExact\ with respect to structural parameters of the underlying graph. We show that the problem is W[1]-hard when parameterized by the feedback vertex number of the underlying graph. 
The \emph{feedback vertex number} of a graph $G$ is the cardinality of a minimum vertex set $X\subseteq V(G)$ such that $G-X$ is a forest. The set $X$ is called a \emph{feedback vertex set}. 
Note that, in contrast to the previous results (\cref{thm:NPhardness}), the reduction we use to obtain the following result does not produce instances with a constant $\Delta$.

\begin{theorem}[$\star$]\label{thm:W1wrtFVS}
    \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

We give a reduction to a variant of \deltaExact\ where the period $\Delta$ is infinite (that is, there are no periods) and we allow $D$ to have infinity entries, meaning that the two respective vertices are not temporally connected. At the end of the proof we show how to obtain the result for a finite $\Delta$ and a matrix $D$ of durations of fastest paths, that only has finite entries.

    Given an instance $H$ of \textsc{Multicolored Clique}, we construct an instance $D$ of \deltaExact\ (with infinity entries and no periods) as follows. To ease the presentation, we first describe the underlying graph $G$ that is implicitly defined by the entries $D_{v,v'}=1$, that is, the pairs of vertices $v,v'$ that should be connected by a temporal path of duration one, meaning that there needs to be an edge connecting the two vertices. Afterwards, we describe the remaining entries of $D$.
    We will construct $G$ using several gadgets.

    \emph{Edge selection gadget.} We first introduce an \emph{edge selection gadget $G_{i,j}$ for color combination $i,j$} with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $4n+1$ vertices each, that is, $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, 
        %u^\ell_n, u^\ell_{n+1},\ldots, u^\ell_{4n-1}, 
        u^\ell_{4n}\}$ for all $\ell\in[m]$. 
        %Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{4n},v_{i,j}^{\star\star}\}$.
